### 2019-11-08 manuscript/manuscript.rmd
upload phylobayes results from choanozoa outgroup

Whole-genome and transcriptomic sequencing is leading to exciting new data matrices for phylogenomic analyses, potentially including hundreds or thousands of genes. Although the number of sampled genes might be different across studies, it is reasonable to expect that the sampled genes should be conserved across animal genomes and most data matrices should share a core set of orthology genes (OGs). Surprisingly, gene composition and matrice overlap across different data matrices has been rarely assessed since most studies relied on their own bioinformatic pipeline for determination of OGs.

@@

The use of site-heterogeneous CAT is an infinte mixture model has long been suggested as the best approach to account for systematic errors, particular for long-branch attraction and saturation. The CAT model often received the best-model fit from cross validation or posterior predictive analyses compared to site-homogeneous models with or without data partitioning. In contrast, a recent study has suggested that Poisson + CAT model did not outperform site-homogeneous with data partitioning or GTR + CAT models and in  cases can infer incorrect topology based on the simulation analyses. The author also raised the issue of the potential overparameterization in CAT model in some cases. Nevertheless，the debate over the site-heterogeneous vs. site-homogeneous models and/or inclusion or exclusion more distantly related outgroups, constitutes the probably one of the most controversial node in the animal tree of life.

@@

One of the challenges of interpreting support for the placement of the animal root across studies is that different programs, software versions, and settings have been used across studies, and phylogenetic analysis decisions have been approached in very different ways. Here we reanalyze the primary matrices from each study under consistent conditions with IQtree under a panel of evolutionary models. We selected this tool because it has greater model flexibility, computational time and accuracy than other tools in maximum likelihood framework.

We first tested a variety of models for each matrix and compared the relative fit of site-homogeneous and site-heterogeneous C models using ModelFinder in IQtree. We found in all cases, C60 + WAG/LG +F +G models fit these matrices bettern than the site-homogeneous models in under BIC criteria and inferred support under the best-fit model from IQtree.  We then analyzed every matrix under a panel of standard site-heterogeneous and site-homogeneous models, including WAG+G, GTR+G and Poisson+C60+F+G. Moreover, we also used the the best-fit model identified above with the removal of all distantly related outgroups and only include Choanozoa as outgroup. With the exception of Moroz2014 matrix (), every analysis conducted herein with different models or outgroup choice strongly supported the ctenophore-sister hypothesis.

It

@@

Site heterogeneity in equilibrium frequency has been a major concern in tests of Ctenophora-sister and Porifera-sister. This has been addressed with CAT models. iqtree provides a new family of C models that also address site heterogeneity. Given the extensive computational cost and concerns about overparameterization of CAT models, we compared iqtree C results to CAT results for a subset of matrices to see if they give consistent results. This would be of technical interest because it would reduce the cost of accommodating compositional heterogeneity in future analyses. For computational efficiency, we first only applied Poisson+CAT models in every matrice with only Choanoflagellata as outgroup. It should be noted that although most Phylobayes runs were converged, the convergence of several large matrices have not been reached after at least a month's running time.

Three matrices with inclusion of Choanoflagellata as outgroup

>>>>

Whole-genome and transcriptomic sequencing is leading to exciting new data matrices for phylogenomic analyses, potentially including hundreds or thousands of genes. Although the number of sampled genes might be different across studies, it is reasonable to expect that the sampled genes should be conserved across animal genomes and most data matrices should share a core set of orthology genes (OGs). Surprisingly, gene composition and matrice overlap across different studies has been rarely assessed, since most studies relied on their own bioinformatic pipeline for determination of OGs.

@@

The use of site-heterogeneous CAT is an infinite mixture model has long been suggested as the best approach to account for systematic errors, particular for long-branch attraction and saturation. The CAT model often received the best-model fit from cross validation or posterior predictive analyses compared to site-homogeneous models with or without data partitioning. In contrast, a recent study has suggested that Poisson + CAT model did not outperform site-homogeneous with data partitioning or GTR + CAT models and in some cases can infer incorrect topology based on the simulation analyses. The author also raised the issue of the potential overparameterization in CAT model in some cases. Nevertheless，the debate over the site-heterogeneous vs. site-homogeneous models and/or inclusion or exclusion more distantly related outgroups, constitutes the probably one of the most controversial node in the animal tree of life.

@@

One of the challenges of interpreting support for the placement of the animal root across studies is that different programs, software versions, and settings have been used across studies, and phylogenetic analysis decisions have been approached in very different ways. Here we reanalyze the primary matrices from each study under consistent conditions with IQtree under a panel of evolutionary models. We selected this tool because it has greater model flexibility, computational time and accuracy than other tools in maximum likelihood (ML) framework.

We first tested a variety of models for each matrix and compared the relative fit of site-homogeneous and site-heterogeneous C models using ModelFinder in IQtree. We found in all cases, C60 + WAG/LG +F +G models fit these matrices better than the site-homogeneous models in under BIC criteria and inferred support under the best-fit model from IQtree.  We didn't include GTR + C models since it needs to estimate over 10,000 free parameters in ModelFinder in some large datasets compared to several hundred in site-homogeneous + C models. then analyzed every matrix under a panel of standard site-heterogeneous and site-homogeneous models, including WAG+G, GTR+G and Poisson+C60+F+G. Moreover, we also used the the best-fit model identified above with the removal of all distantly related outgroups and only include Choanozoa as outgroup.

With the exception of Moroz2014 matrix supported Placozoa-sister, every analysis conducted herein with different models or outgroup choice strongly supported the ctenophore-sister hypothesis under maximum likelihood framework in IQtree.

The site-heterogeneous CAT model has been introduced in PhyloBayes and it is a non-parametric modeling based on a Dirichlet process mixture, which is often used in Bayesian studies (Lartillot and Philippe, 2004). It was designed to better accommodate data saturation or tackle long-branch attraction for deep phylogenetic question but only used in Bayesian inference (BI). In contrast, C models is 10 to 60-profile mixture models as variants of the CAT model can be used in both ML or BI framework. Thus, it is interesting to compare results between C model to CAT models since it has never been used in any matrices analyzed here.

@@

Site heterogeneity in equilibrium frequency has been a major concern in tests of Ctenophora-sister and Porifera-sister. This has been addressed with CAT models. iqtree provides a new family of C models that also address site heterogeneity. Given the extensive computational cost and concerns about overparameterization of CAT models, we compared iqtree C results to CAT results for a subset of matrices to see if they give consistent results. This would be of technical interest because it would reduce the cost of accommodating compositional heterogeneity in future analyses. For computational efficiency, we first only applied Poisson+CAT models in every matrice with only Choanoflagellata as outgroup. It should be noted that most Phylobayes runs were converged, although several large matrices have not reached convergence after at least a month's computational time.

Interestingly, we found a mixed support when using Poisson-CAT model in Choanimalia matrices.  Consistent with previous study, we recovered Porifera-sister in Philippe2008, Ryan2013, Nosenko2013 matrices with only. Moreover, we also found a strong support of Porifera-sister in that Whelan2017 matrices. However, with the exception of Moroz2013, we found strong support for Ctenophora-sister in all other matrices.
